---
title: "Data Analysis"
author: "Ricardo Beltran"
date: 2021-06-08T21:13:14-05:01
categories: ["R"]
tags: ["R Markdown", "plot", "regression"]
draft: FALSE
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(collapse = TRUE)
```

```{r libraries, echo=FALSE, include=FALSE}
library(tidyverse)
library(lubridate)
```

```{r read_data}

game_data <- read.csv(here:::here("Data", "games_details.csv"), stringsAsFactors = FALSE)

```

# Intro:

This page will take a look at and analyze several metrics from our NBA data set.

# Preprocessing
```{r}

game_data <- game_data %>%
  mutate(
    TEAM_ID = as.factor(TEAM_ID),
    TEAM_ABBREVIATION = as.factor(TEAM_ABBREVIATION),
    TEAM_CITY = as.factor(TEAM_CITY),
    PLAYER_ID = as.factor(PLAYER_ID),
    START_POSITION = case_when(
      (START_POSITION == "C" |
      START_POSITION == "F" |
      START_POSITION == "G") ~ START_POSITION,
      TRUE ~ "Didn't Start"
    ), 
    STARTER_STATUS = case_when(
      (START_POSITION == "C" |
      START_POSITION == "F" |
      START_POSITION == "G") ~ "STARTER",
      TRUE ~ "NON-STARTER"
    ),
    STARTER_STATUS = as.factor(STARTER_STATUS),
    START_POSITION = as.factor(START_POSITION)
  )
    
game_data[is.na(game_data)] <- 0


```

# What's in a Name?

For our early analysis, I'm interested to see if there are any common names among the players in the data set.

```{r}

names <- game_data %>%
  group_by(PLAYER_NAME) %>%
  summarise(junk = 1) %>%
  select(PLAYER_NAME) %>%
  pull() %>% 
  str_split(" ") %>%
  unlist()

names <- names[names != "Jr." & names != "jr."]
  
as.data.frame(table(names)) %>%
  top_n(10, Freq)

# This included some ties for those with a freq of 2
```
Here we found the 10 most common names from among the NBA players in the data set. It appears that names like Johnson, Chris and Williams were the most popular. 

Fun Fact: There have been multiple ALL-STARS and even HALL OF FAMERS with those names.


## Starters vs Non-Starters

In the game of basketball, while it's not always the case, the best players usually start the game, hence their title of starters. I'm interested to see just how much of a performance gap there is between these two groups.

```{r}

starter_grouped <- game_data %>%
  group_by(START_POSITION) %>%
  summarise(avg_points = mean(PTS), 
            avg_rebounds = mean(REB), 
            avg_assists = mean(AST))

ggplot(starter_grouped, aes(x=START_POSITION, y=avg_points )) + 
  geom_bar(stat = "identity") +
  xlab("Starter Status") + 
  ylab("Average Points") + 
  ggtitle("Average Points for Different Starter Positions") + 
  theme_bw() +
  geom_col(fill = "darkorchid") 
```

```{r}
ggplot(starter_grouped, aes(x=START_POSITION, y=avg_rebounds )) + 
  geom_bar(stat = "identity") +
  xlab("Starter Status") + 
  ylab("Average Rebounds") + 
  ggtitle("Average Rebounds for Different Starter Positions") + 
  theme_bw() +
  geom_col(fill = "darkorchid")
```

```{r}
ggplot(starter_grouped, aes(x=START_POSITION, y=avg_assists )) + 
  geom_bar(stat = "identity") +
  xlab("Starter Status") + 
  ylab("Average Assists") + 
  ggtitle("Average Assists for Different Starter Positions") + 
  theme_bw() +
  geom_col(fill = "darkorchid")

```
From these graphs, it's clear that Non-Starters did not produce the same volume of points, assists, or rebounds, which are the main metrics in basketball. It also makes sense with today's game why Guards would be making more points than centers, and centers more rebounds than any other group. 

However, perhaps starters play for longer and can accumulate more of these, so a better metric would be their percentage performances for other metrics, such as normal fieldgoals, field goals, and three pointers. 

# Performance Percentages
```{r}

starter_grouped <- game_data %>%
  group_by(START_POSITION) %>%
  summarise(avg_fg_pct = mean(FG_PCT), 
            avg_fg3_pct = mean(FG3_PCT), 
            avg_ft_pct = mean(FT_PCT))

ggplot(starter_grouped, aes(x=START_POSITION, y=avg_fg_pct)) + 
  geom_bar(stat = "identity") +
  xlab("Starter Status") + 
  ylab("Average Field Goal Percentage") + 
  ggtitle("Mean Field Goal Percentage for Different Starter Positions") + 
  theme_bw() +
  geom_col(fill = "darkorchid") 

ggplot(starter_grouped, aes(x=START_POSITION, y=avg_fg3_pct  )) + 
  geom_bar(stat = "identity") +
  xlab("Starter Status") + 
  ylab("Mean 3-Pointer Percentage") + 
  ggtitle("Mean 3-Pointer Percentage for Different Starter Positions") + 
  theme_bw() +
  geom_col(fill = "darkorchid")

ggplot(starter_grouped, aes(x=START_POSITION, y=avg_ft_pct )) + 
  geom_bar(stat = "identity") +
  xlab("Starter Status") + 
  ylab("Mean Free Throw Percentage") + 
  ggtitle("Mean Free Throw Percentage for Different Starter Positions") + 
  theme_bw() +
  geom_col(fill = "darkorchid")

```
Here we find that while, overall, the Starters still outperformed the non-starters, this was not the case with the 3-Pointers. If the data was present it would be helpful if we could see what position the non starters were.

You can also embed plots. See Figure \@ref(fig:pie) for example:



```{r}
model <- game_data %>%
  lm(PTS ~ FGA, data = .)

game_data %>%
  ggplot(aes(x = FGA, y = PTS)) +
  geom_point() +
  stat_smooth(method = "lm")

broom::tidy(model)
```

From the metrics given and the visible pattern, it's clear that the more field goals are attempted, the more points one makes generally speaking. This might seem a bit trivial, but if only certain players understood that it's ok to shoot more. You miss 100% of the shots you don't take after all!

```{r}

list <- lapply(strsplit(game_data$MIN, ":")[[1]], function(x) 
       scan(text = x, what = numeric(), sep=","))

list %>%
  unlist() %>%
  mean()
```
Interestingly enough, the average NBA player here plays about 35 mins, which is quite a bit considering that a regulation game is 48 minutes.

