---
title: "Getting to Know the Data Set"
author: "Ricardo Beltran"
date: 2021-06-08T21:13:14-05:10
categories: ["R"]
tags: ["R Markdown", "plot", "regression"]
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(collapse = TRUE)
```

# R Markdown

The data set we'll be looking at contains data about several NBA, in-game statistics. These include metrics such as:
- GAME_ID   

- TEAM_ID 

- TEAM_ABBREVIATION  

- TEAM_CITY

- PLAYER_ID   

- PLAYER_NAME

- START_POSITION COMMENT   

- MIN : Minutes played

- FGM : Field Goals Made

- FGA : Field Goals Attempted

- FG_PCT : Field Goal Percentage

- FTM : Free throws made

- FTA : Free throws attempted

- FT_PCT : Free throw percentage

AND OTHERS!

It will allow us to take a better look at players, their stats, and several relationships between the origin, build, team, playstyle of individuals and groups and how well or poorly they performed. 

For more details on this data set and other like it, check out the following link <https://www.kaggle.com/nathanlauga/nba-games?select=games_details.csv>.

The Data Comes Packaged as Such:
```{r read_data}

game_data <- read.csv(here:::here("Data", "games_details.csv"), stringsAsFactors = FALSE)

```

```{r}

game_data
```
